---
title: |
  <div class="custom-title-block" style="font-size: 1.2em;">
    <span style="color:#000000;">Replication of below article's Data and Visualizations</span><br>
    <span style="color:#333333; font-size:0.8em;">"These Are the U.S. Universities Most Dependent on International Students"</span><br>
    <span style="color:#666666; font-size:0.7em;">
      By <a href="https://www.nytimes.com/2025/05/23/upshot/harvard-trump-international-students.html" 
      target="_blank" style="color:#000000; text-decoration:underline;">Emily Badger, Aatish Bhatia, Asmaa Elkeurtim Steven Rich and Ethan Singer"</a><br>
    </span><br>
    <span style="font-size:0.8em; color:#333333; white-space: nowrap">
      Karim K. Kardous
      <a href='mailto:kardouskarim@gmail.com' style='margin-left: 9px; font-size: 0.9em;'>
        <i class='bi bi-envelope'></i>
      </a>
      <a href='https://github.com/kkardousk' style='margin-left: 5px; font-size: 0.9em;'>
        <i class='bi bi-github'></i>
      </a>
    </span>
  </div>
format:  
  html:
    toc: true
    toc-depth: 4
    toc-expand: true
    toc-title: 'Jump To'
    number-depth: 2
    fig-format: retina
    fig-dpi: 300
    code-link: true # requires both downlit and xml2 to be downloaded
    code-fold: true
    code-summary: '<i class="bi-code-slash"></i> Show the code'
    # code-overflow: wrap
    code-tools:
      toggle: true  # adds "Show All / Hide All"; also allows for all code copy (at once as quarto doc)
    css: styles.css
    highlight-style: github-dark
    df-print: paged
    page-layout: article
    embed-resources: true
    smooth-scroll: true
    link-external-icon: false
    link-external-newwindow: true
    fontsize: 1.1em
    linestretch: 0
    linespace: 0
    html-math-method: katex
    linkcolor: '#D35400'
execute:
  echo: true
  warning: false
  message: false
  info: false
  cache: false
  freeze: auto
editor: visual
---

## **Initial Setup**

::: text-justify
I start by configuring the Python environment using `{reticulate}`, specifying a custom virtual environment and ensuring required packages (matplotlib and pillow) are installed. This setup step ensures Python and R work seamlessly together; granted 'seamlessly' might be over-reaching, but at least for this case, fairly commonly used modules, it was the case.
:::

```{r}
#| echo: true
#| message: false
library(reticulate)
invisible(
  capture.output({
    Sys.setenv(RETICULATE_PYTHON = "py_venv/bin/python3.9")
    reticulate::use_python("py_venv/bin/python3.9", required = TRUE)
    })
)

invisible(
  capture.output({
  use_python("py_venv/bin/python3.9", required = TRUE)
  py_config()
  py_install(c('playwright', 'beautifulsoup4', 'pandas', 'plotnine'), envname = "py_venv", pip = TRUE)
  })
)
```

```{r}
system("py_venv/bin/python -m playwright install chromium")
```


